Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-07-22 exp1/iswc2013.tex
Martins suggestions
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@604 43ef1190-90fd-4dd7-a769-d0121f02b80e

One important aspect of science is the structured approach to falsify
hypotheses, and thus is important that assertions about performance is
formulated as hypotheses that can be falsified.

Thus, attacking the complexity problem by eliminating parameters we
think influence the outcome for simplicity is untenable as scientific
practice. We must instead tackle this problem by enhancing our
capability of testing complex systems and to identify situations where
our assumptions are flawed. Such constraints can only be made when we
have very good reason to suspect that a certain parameter does not
have any influence.

@@

A literature survey has not revealed any direct prior art, however, the
general approach has been well established, not only in statistics. A
relatively well cited textbook is \cite{citeulike:5190414} but we have
not found the parts discussed in the present paper to be widely
adopted. Recently, a comprehensive text on experimental methods have
been published in
\cite{Springer-2010-Experimental-Methods-for-the-Analysis-of-Optimization-Algorithms}
demonstrating the broad applicability of the methodology we employ.
We have chosen SPARQL Endpoint evaluation as an example in this study
since SPARQL Federation is our main interest, and to turn to
statistical standard texts \cite{wu2009experiments} for this study.

@@

Finally, we acknowledge the efforts of the Linked Data Benchmark
Council and the SEALS project. However, they appear to focus on
industrial benchmarking rather than scientific evaluations.

@@

Common experiments include comparing several different
implementations. However, we focus on a slightly different problem: We
would like to compare the same implementation before and after some
change has been made, typically with the intention to enhance the
performance of the system. There are two reasons for this: One is that
there are currently many SPARQL implementations available, also Free
Software ones that can be modified by anyone, and the other is to
constrain the scope of the study to comparing two different things.

@@

These fragments has been carefully designed for illustrative utility,
as well as suitable selectivity. They are not themselves important,
they only serve to illustrate certain examples of possible
factors. When the experiment is run, they are combined to yield a
query, which is then sent to a SPARQL endpoint. The SPARQL endpoint is
itself set up using 4store (see \cite{harris20094store}) version
1.1.5. The ``Machine'' factor specifies a hostname, whereas
``TripleC'' and ``Implement'' specifies a port number. On each host,
four 4store instances runs independently on different ports.

>>>>

One important aspect of science is the structured approach to falsify
hypotheses, and thus is important that assertions about performance
are formulated as hypotheses that can be falsified.

Thus, attacking the complexity problem by eliminating parameters we
think influence the outcome for simplicity is untenable as scientific
practice. We must instead tackle this problem by enhancing our
capability of testing complex systems and to identify situations where
our assumptions are flawed. Such constraints can only be made when we
have highly confident that a certain parameter does not have any
influence.

@@

A literature survey has not revealed any direct prior art, however, the
general approach has been well established, not only in statistics. A
relatively well cited textbook is \cite{citeulike:5190414} but we have
not found the parts discussed in the present paper to be widely
adopted. Recently, a comprehensive text on experimental methods has
been published in
\cite{Springer-2010-Experimental-Methods-for-the-Analysis-of-Optimization-Algorithms}
demonstrating the broad applicability of the methodology we employ.
We have chosen SPARQL Endpoint evaluation as an example in this study
since SPARQL Federation is our main interest, and to turn to
statistical standard texts \cite{wu2009experiments} for this study.

@@

Finally, we acknowledge the efforts of the Linked Data Benchmark
Council and the SEALS project. However, they appear to focus on
industrial benchmarking rather than scientific evaluations.
% TODO: skrive om til å snakke om use-case driven

@@

Common experiments include comparing several different
implementations. However, we focus on a slightly different problem: We
would like to compare the same implementation before and after some
change has been made, typically with the intention to enhance the
performance of the system. There are two reasons for this: One is that
there are currently many SPARQL implementations available, also Free
Software ones that can be modified by anyone, and the other is to
constrain the scope of the study to comparing just two different
things.

@@

These fragments has been carefully designed for illustrative utility,
as well as suitable selectivity. They are not themselves important,
they only serve to illustrate certain examples of possible
factors. When the experiment is run, they are combined to yield a
query, which is then sent to a SPARQL endpoint. The SPARQL endpoint is
itself set up using 4store (see \cite{harris20094store}) version
1.1.5. The ``Machine'' factor specifies a hostname, whereas
``TripleC'' and ``Implement'' specifies a port number. On each host,
four 4store instances runs independently on different ports. 
% TODO: process waiting.
